Design and Implementation of a Multithreaded Virtual Machine for Executing Linear Logic Programs

### 2014-08-07 experiments.tex
Update.
Fixed.
ricroc's changes.

In Table~\ref{tbl:comp_bp}, we compare LM's Belief Propagation (BP)
program, a machine learning algorithm to denoise images, against a
sequential C, Python and GraphLab~\cite{GraphLab2010} version of the
algorithm. GraphLab is a parallel C++ library used to solve
graph-based problems in machine learning. C and GraphLab perform about
the same since both use C/C++. Python runs very slowly since it is a
dynamic programming language and BP has many mathematical
computations. We should note, however, that LM's version uses some
external functions written in C++ in order to improve execution time,
therefore the comparison is not totally fair.

\begin{table}[ht]
\centering
{\begin{tabular}{c|c|c|c}
\textbf{Problem} & \multicolumn{3}{c}{\textbf{System}} \\
\textbf{Size} & \textbf{C} & \textbf{Python} & \textbf{GraphLab} \\
\hline\hline
\textbf{10}  & 0.67 & 0.03 & 1.00 \\
\textbf{50}  & 1.77 & 0.04 & 1.73 \\
\textbf{200} & 1.99 & 0.05 & 1.79 \\
\textbf{400} & 2.00 & 0.04 & 1.80 \\
\end{tabular}}
\caption{Comparing the absolute execution times (LM/System) for the
  Belief Propagation program}
\label{tbl:comp_bp}
\end{table}

@@

\begin{figure*}[ht]
\centering
\subfigure[PageRank using a graph of web pages with around 12,000 nodes and 292,000 edges]
   {\label{exp:pagerank}\includegraphics[width=\plotsize]{figures/results-pagerank-search-engines.pdf}}
\hspace{0.5cm}
\subfigure[GGC using a random graph with 2,000 nodes and 600,000 edges]
   {\label{exp:ggc}\includegraphics[width=\plotsize]{figures/results-ggc.pdf}}
\caption{Experimental results for the GGC and PageRank algorithms}
\end{figure*}

@@

\begin{figure*}[ht]
\centering
\subfigure[N-Queens program (13x13~board)]
   {\label{exp:13queens}\includegraphics[width=\plotsize]{figures/results-13queens.pdf}}
\hspace{0.5cm}
\subfigure[BP (400x400~image)]
   {\label{exp:bp}\includegraphics[width=\plotsize]{figures/results-bp.pdf}}
\caption{Experimental results for the N-Queens and Belief Propagation programs}
\end{figure*}

@@

In this section, we present the amount of \textbf{Memory} used by the VM after the
programs presented in the previous section complete. We also count the number of
\textbf{Facts} stored in the database. This serves as an indication of
how much memory is needed in terms of the number of facts stored in the database.
Results are shown in Table~\ref{tbl:memory}.

\begin{table}[ht]
\centering
{\begin{tabular}{c|c|c|c}
\textbf{Program} & \textbf{Facts} & \textbf{Memory} & \textbf{Per Fact} \\
\hline\hline
PageRank & 1,180,603 & 203,832 KB & 176 bytes \\
GGC & 2,363,536 & 292,682 KB & 127 bytes \\
SD & 502,347 & 45,387 KB & 92 bytes \\
MiniMax & 3 & 886,443 KB & 295,481 KB \\
N-Queens & 74,557 & 55,408 KB & 760 bytes \\
BP & 2,235,200 & 617,417 KB & 283 bytes \\
\hline\hline
\multicolumn{3}{l}{Average (W/ MiniMax) } & 288 bytes \\
\end{tabular}}
\caption{Memory usage of programs after completion using 1 thread.}
\label{tbl:memory}
\end{table}

@@

PageRank, GGC and SD show a low memory usage per fact. This makes sense since the predicates
used in those programs are relatively simple with only a few arguments.
This changes with the N-Queens problem where the predicates have lists representing the valid
positions of the queens in the board. Storing lists is far more expensive than storing integral
values such as integers or floating point numbers. The same thing happens with BP because
the belief values are stored as lists of floating point numbers.

>>>>

In Table~\ref{tbl:comp_bp}, we compare LM's Belief Propagation (BP)
program, a machine learning algorithm to denoise images, against a
sequential C, Python and GraphLab~\cite{GraphLab2010} version of the
algorithm. GraphLab is a parallel C++ library used to solve
graph-based problems in machine learning. C and GraphLab perform about
the same since they are both compiled to machine code, although
the GraphLab version is highly optimized to run on multicore architectures.
Python runs very slowly since it is a
dynamic programming language and BP has many mathematical
computations. We should note, however, that LM's version uses some
external functions written in C++ in order to improve execution time,
therefore the comparison is not totally fair.

\begin{table}[ht]
\centering
{\begin{tabular}{c|c|c|c}
\textbf{Problem} & \multicolumn{3}{c}{\textbf{System}} \\
\textbf{Size} & \textbf{C} & \textbf{Python} & \textbf{GraphLab} \\
\hline\hline
\textbf{10}  & 1.00 & 0.03 & 1.00 \\
\textbf{50}  & 1.77 & 0.04 & 1.73 \\
\textbf{200} & 1.99 & 0.05 & 1.79 \\
\textbf{400} & 2.00 & 0.04 & 1.80 \\
\end{tabular}}
\caption{Comparing the absolute execution times (LM/System) for the
  Belief Propagation program}
\label{tbl:comp_bp}
\end{table}

@@

\begin{figure*}[ht]
\centering
\subfigure[PageRank using a graph of web pages with around 12,000 nodes and 292,000 edges]
   {\label{exp:pagerank}\includegraphics[width=\plotsize]{figures/results-pagerank-search-engines.pdf}}
\hspace{0.5cm}
\subfigure[GGC using a random graph with 2,000 nodes and 600,000 edges]
   {\label{exp:ggc}\includegraphics[width=\plotsize]{figures/results-ggc.pdf}}
\caption{Experimental results for the PageRank and GGC  algorithms}
\end{figure*}

@@

\begin{figure*}[ht]
\centering
\subfigure[N-Queens program (13x13~board)]
   {\label{exp:13queens}\includegraphics[width=\plotsize]{figures/results-13queens.pdf}}
\hspace{0.5cm}
\subfigure[Belief Propagation (400x400~image)]
   {\label{exp:bp}\includegraphics[width=\plotsize]{figures/results-bp.pdf}}
\caption{Experimental results for the N-Queens and Belief Propagation programs}
\end{figure*}

@@

In this section, we present the amount of memory used by the VM after the
programs presented in the previous section complete. We also count the number of
facts stored in the database. This serves as an indication of
how much memory is needed in terms of the number of facts stored in the database.
Results are shown in Table~\ref{tbl:memory}.

\begin{table}[ht]
\centering
{\begin{tabular}{c|c|c|c}
\textbf{Program} & \textbf{Facts} & \textbf{Memory} & \textbf{Per Fact} \\
\hline\hline
PageRank & 1,180,603 & 203,832 KB & 176 bytes \\
GGC & 2,363,536 & 292,682 KB & 127 bytes \\
SD & 502,347 & 45,387 KB & 92 bytes \\
MiniMax & 3 & 886,443 KB & 295,481 KB \\
N-Queens & 74,557 & 55,408 KB & 760 bytes \\
BP & 2,235,200 & 617,417 KB & 283 bytes \\
\hline\hline
\multicolumn{3}{l}{Average (without MiniMax) } & 288 bytes \\
\end{tabular}}
\caption{Memory usage of programs after completion using 1 thread.}
\label{tbl:memory}
\end{table}

@@

PageRank, GGC and SD show a low memory usage per fact. This makes sense since the predicates
used in those programs are relatively simple with only a few arguments.
This changes with the N-Queens problem where the predicates have lists representing the valid
positions of the queens in the board. Storing lists is far more expensive than storing integral
values such as integers or floating point numbers. The same happens with BP because
the belief values are stored as lists of floating point numbers.

### 2014-08-07 introduction.tex
ricroc's changes.

\iffalse
Logic programming languages split into two main
groups: \emph{backwards-chaining} and \emph{forward-chaining}. In a
backwards-chaining logic programming language, programs are composed
of a set of rules that can be activated by inputting a query. Given a
query $q(\hat{x})$, an interpreter will work backwards by matching
$q(\hat{x})$ against the head of a rule. If found, the interpreter
will then try to match the body of the rule, recursively, until it
finds the program axioms (rules without body). If the search procedure
succeeds, the interpreter finds a valid substitution for the $\hat{x}$
variables. A popular backwards-chaining programming language is
Prolog~\cite{Colmerauer:1993:BP:154766.155362}, which has been a
productive research language for executing logic programs in
parallel. Researchers took advantage of Prolog's non-determinism to
evaluate subgoals in parallel with models such
as \emph{And-parallelism}
and \emph{Or-parallelism}~\cite{Gupta:2001:PEP:504083.504085}.
\fi

In forward-chaining logic programming, programs start with
a database of facts (filled with the program's axioms) and a set of
logical rules. The database of facts is then used to fire the
program's rules and derive new facts that are then added to the
database. This process is repeated recursively until the database
reaches \emph{quiescence} and no more information can be derived from
the program. A popular forward-chaining programming language is
Datalog~\cite{Ramakrishnan93asurvey}.

@@

In this paper, we present the design and implementation of the LM virtual machine
and compiler.
The LM virtual machine was designed from scratch to run LM programs on
multicore machines in an efficient manner\footnote{Source code available at \url{http://github.com/flavioc/meld}}
The virtual machine is multithreaded and executes byte-code that is generated
by the LM compiler. To test our language and virtual machine we have implemented several graph algorithms, search
algorithms and machine learning algorithms, including: belief
propagation~\cite{Gonzalez+al:aistats09paraml}, belief propagation
with residual splash~\cite{Gonzalez+al:aistats09paraml}, PageRank,
graph coloring, N-Queens, shortest path, diameter estimation, map
reduce, quick-sort, neural network training, minimax, etc.

@@

This paper is organized as follows. First, we briefly introduce the LM
language. Then, we present an overview of the virtual machine and
describe in more detail the code organization, thread management, rule
execution and database organization. Finally, we present our experiments
and outline some conclusions.

>>>>

Logic programming split into two main groups:
\emph{backwards-chaining} and \emph{forward-chaining} languages. In a
backwards-chaining programming language, programs are composed of a
set of rules that can be activated by inputting a query. Given a query
$q(\hat{x})$, an interpreter will work backwards by matching
$q(\hat{x})$ against the head of a rule. If found, the interpreter
will then try to match the body of the rule, recursively, until it
finds the program axioms (rules without body). If the search procedure
succeeds, the interpreter finds a valid substitution for the $\hat{x}$
variables. A popular backwards-chaining programming language is
Prolog~\cite{Colmerauer:1993:BP:154766.155362}, which has been a
productive research language for executing logic programs in
parallel. Researchers took advantage of Prolog's non-determinism to
evaluate subgoals in parallel with models such as
\emph{And-parallelism} and
\emph{Or-parallelism}~\cite{Gupta:2001:PEP:504083.504085}.

In a forward-chaining logic programming language, programs start with
a database of facts (filled with the program's axioms) and a set of
logical rules. The database of facts is then used to fire the
program's rules and derive new facts that are then added to the
database. This process is repeated recursively until the database
reaches \emph{quiescence} and no more information can be derived from
the program. A popular forward-chaining programming language is
Datalog~\cite{Ramakrishnan93asurvey}.

@@

In this paper, we present the design and implementation of the LM virtual machine
and compiler.
The LM virtual machine was designed from scratch to run LM programs on
multicore machines in an efficient manner\footnote{Source code available at \url{http://github.com/flavioc/meld}}.
The virtual machine is multithreaded and executes byte-code that is generated
by the LM compiler. To test our language and virtual machine we have implemented several graph algorithms, search
algorithms and machine learning algorithms, including: belief
propagation~\cite{Gonzalez+al:aistats09paraml}, belief propagation
with residual splash~\cite{Gonzalez+al:aistats09paraml}, PageRank,
graph coloring, N-Queens, shortest path, diameter estimation, map
reduce, quick-sort, neural network training, minimax, etc.

@@

This remainder of the paper is organized as follows. First, we briefly
introduce the LM language. Then, we present an overview of the virtual
machine and describe in more detail the code organization, thread
management, rule execution and database organization. Finally, we
present our experiments and outline some conclusions.

### 2014-08-07 language.tex
More data
ricroc's changes.

\begin{figure}[t]
{\footnotesize
\begin{Verbatim}[numbers=right]
type left(node, node).
type right(node, node).
type linear value(node, int, string).
type linear replace(node, int, string).

@@

The algorithm uses three rules for the three cases of updating a key's
value: the first rule performs the update (lines 6-9); the second rule
recursively picks the left branch for the update operation (lines
11-16); and the third rule picks the right branch (lines 18-23). The
initial axioms are presented in lines 26-33 and they describe the
initial binary tree configuration, including keys and values.
Finally, with the \texttt{replace(@0, 6, x)} axiom instantiated at the
root node $@0$, we intend to change the value of key 6 to value
`$x$'. Note that when writing rules or axioms, persistent facts are
preceded with a \texttt{!}.

@@

Comprehensions and aggregates are logically justified by the underlying
proof system of the language. Our proof system is extended with greatest
fixed points~\cite{Baelde:2012:LGF:2071368.2071370}, which allows us to describe
recursive definitions such as comprehensions or aggregates.

>>>>

\begin{figure}[ht]
{\footnotesize
\begin{Verbatim}[numbers=right]
type left(node, node).
type right(node, node).
type linear value(node, int, string).
type linear replace(node, int, string).

@@

The algorithm uses three rules for the three cases of updating a key's
value: the first rule performs the update (lines 6-9); the second rule
recursively picks the left branch for the update operation (lines
11-16); and the third rule picks the right branch (lines 18-23). The
initial axioms are presented in lines 26-33 and they describe the
initial binary tree configuration, including keys and values.
Finally, with the \texttt{replace(@0, 6, x)} axiom instantiated at the
root node \texttt{@0} (line 36), we intend to change the value of key
6 to value \texttt{x}. Note that when writing rules or axioms,
persistent facts are preceded with a `\texttt{!}'.

@@

Comprehensions and aggregates are logically justified by the underlying
proof system of the language. Our proof system is extended with greatest
fixed points~\cite{Baelde:2012:LGF:2071368.2071370}, which allows us to describe
recursive definitions such as comprehensions or aggregates.
A more detailed description is found in~\cite{cruz-iclp14}.

### 2014-08-07 paper.tex
ricroc's changes.

\documentclass{acm_proc_article-sp}
\usepackage[utf8]{inputenc}
\usepackage{epsfig}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{subfigure}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{balance}
\usepackage{listings}
\usepackage{multicol}
\lstset{language=C, basicstyle=\ttfamily, breaklines=true, frame=lines}
\input{fp-macros}

@@

@@

\numberofauthors{3}
\author{
  \alignauthor Flavio Cruz \\
    \affaddr{Carnegie Mellon University}
    \affaddr{Pittsburgh, PA 15213, USA}\\
    \email{fmfernan@cs.cmu.edu} \\
  \alignauthor Ricardo Rocha \\
    \affaddr{CRACS \& INESC TEC}\\
    \affaddr{University of Porto}\\
    \affaddr{Rua Campo Alegre 1021/1055}\\
    \affaddr{4169-007 Porto, Portugal}\\
    \email{ricroc@dcc.fc.up.pt}
  \alignauthor Seth Copen Goldstein \\
    \affaddr{Carnegie Mellon University}
    \affaddr{Pittsburgh, PA 15213, USA}\\
    \email{seth@cs.cmu.edu} \\
}

@@

\input{abstract}
\input{introduction}
\input{language}
\input{virtual_machine}
\input{experiments}
\input{related_work}
\input{conclusions}

@@

\bibliographystyle{abbrv}
\bibliography{refs}
\end{document}

>>>>

\documentclass{sigplanconf}
\usepackage[utf8]{inputenc}
\usepackage{epsfig}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{subfigure}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{balance}
\usepackage{listings}
\usepackage{multicol}
\lstset{language=C, basicstyle=\ttfamily, breaklines=true, frame=lines}
\input{fp-macros}

@@

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{PPDP '14}{September 08--10, 2014, Canterbury, United Kingdom}
\copyrightyear{2014} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.
\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright
%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

@@

\authorinfo{Flavio Cruz}
           {Carnegie Mellon University\\Pittsburgh, PA 15213, USA}
           {fmfernan@cs.cmu.edu}

\authorinfo{Ricardo Rocha}
           {CRACS \& INESC TEC\\University of Porto\\Rua Campo Alegre 1021/1055\\4169-007 Porto, Portugal}
           {ricroc@dcc.fc.up.pt}

\authorinfo{Seth Copen Goldstein}
           {Carnegie Mellon University\\Pittsburgh, PA 15213, USA}
           {seth@cs.cmu.edu}

@@

\input{abstract}
\input{introduction}
\input{language}
\input{virtual_machine}
\input{experiments}
\input{related_work}
\input{conclusions}
\input{ack}

@@

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.
%\bibliography{refs}

\begin{thebibliography}{35}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Ali and Karlsson(1990)]{Ali:1990fk}
K.~Ali and R.~Karlsson.
\newblock Full prolog and scheduling or-parallelism in muse.
\newblock \emph{International Journal of Parallel Programming}, 19\penalty0
  (6):\penalty0 445--475, 1990.
\newblock ISSN 0885-7458.
\newblock \doi{10.1007/BF01397627}.
\newblock URL \url{http://dx.doi.org/10.1007/BF01397627}.

\bibitem[Ali(1988)]{Ali88}
K.~A.~M. Ali.
\newblock Or-parallel execution of prolog on bc-machine.
\newblock In R.~A. Kowalski and K.~A. Bowen, editors, \emph{ICLP/SLP}, pages
  1531--1545. MIT Press, 1988.
\newblock ISBN 0-262-61056-6.
\newblock URL \url{http://dblp.uni-trier.de/db/conf/iclp/iclp88.html#Ali88}.

\bibitem[Ashley-Rollman et~al.(2007)Ashley-Rollman, Rosa, Srinivasa, Pillai,
  Goldstein, and Campbell]{ashley-rollman-derosa-iros07wksp}
M.~P. Ashley-Rollman, M.~D. Rosa, S.~S. Srinivasa, P.~Pillai, S.~C. Goldstein,
  and J.~D. Campbell.
\newblock Declarative programming for modular robots.
\newblock In \emph{Workshop on Self-Reconfigurable Robots/Systems and
  Applications at IROS 2007}, 2007.

\bibitem[Ashley-Rollman et~al.(2009)Ashley-Rollman, Lee, Goldstein, Pillai, and
  Campbell]{ashley-rollman-iclp09}
M.~P. Ashley-Rollman, P.~Lee, S.~C. Goldstein, P.~Pillai, and J.~D. Campbell.
\newblock A language for large ensembles of independently executing nodes.
\newblock In \emph{International Conference on Logic Programming (ICLP)}, 2009.

\bibitem[Baelde(2012)]{Baelde:2012:LGF:2071368.2071370}
D.~Baelde.
\newblock Least and greatest fixed points in linear logic.
\newblock \emph{ACM Transactions on Computational Logic}, 13\penalty0
  (1):\penalty0 1--44, 2012.

\bibitem[Betz and Fr{\"u}hwirth(2005)]{Betz:2005kx}
H.~Betz and T.~Fr{\"u}hwirth.
\newblock A linear-logic semantics for constraint handling rules.
\newblock In \emph{Principles and Practice of Constraint Programming - CP
  2005}, volume 3709 of \emph{Lecture Notes in Computer Science}, pages
  137--151. 2005.

\bibitem[Butler et~al.(1988)Butler, Disz, Lusk, Olson, Overbeek, and
  Stevens]{ButlerDLOOS88}
R.~Butler, T.~Disz, E.~L. Lusk, R.~Olson, R.~A. Overbeek, and R.~L. Stevens.
\newblock Scheduling or-parallelism: An argonne perspective.
\newblock In R.~A. Kowalski and K.~A. Bowen, editors, \emph{ICLP/SLP}, pages
  1590--1605. MIT Press, 1988.
\newblock ISBN 0-262-61056-6.
\newblock URL
  \url{http://dblp.uni-trier.de/db/conf/iclp/iclp88.html#ButlerDLOOS88}.

\bibitem[Colmerauer and Roussel(1993)]{Colmerauer:1993:BP:154766.155362}
A.~Colmerauer and P.~Roussel.
\newblock The birth of prolog.
\newblock In \emph{The Second ACM SIGPLAN Conference on History of Programming
  Languages}, pages 37--52, New York, NY, USA, 1993.

\bibitem[Costa et~al.(2011)Costa, Damas, and
  Rocha]{DBLP:journals/corr/abs-1102-3896}
V.~S. Costa, L.~Damas, and R.~Rocha.
\newblock The yap prolog system.
\newblock \emph{CoRR}, abs/1102.3896, 2011.

\bibitem[Cruz et~al.(2014)Cruz, Rocha, Goldstein, and Pfenning]{cruz-iclp14}
F.~Cruz, R.~Rocha, S.~Goldstein, and F.~Pfenning.
\newblock {A Linear Logic Programming Language for Concurrent Programming over
  Graph Structures}.
\newblock \emph{Journal of Theory and Practice of Logic Programming, 30th
  International Conference on Logic Programming (ICLP 2014), Special Issue},
  14\penalty0 (4 \& 5):\penalty0 493--507, July 2014.

\bibitem[Culler et~al.(1993)Culler, Goldstein, Schauser, and von
  Eicken]{CullerGSvE93}
D.~E. Culler, S.~C. Goldstein, K.~E. Schauser, and T.~von Eicken.
\newblock {TAM --- a compiler controlled threaded abstract machine}.
\newblock \emph{Journal of Parallel and Distributed Computing}, 18:\penalty0
  347--370, July 1993.
\newblock URL \url{http://www.cs.cmu.edu/~seth/papers/CullerGSvE93.pdf}.

\bibitem[Girard(1987)]{girard-87}
J.-Y. Girard.
\newblock Linear logic.
\newblock \emph{Theoretical Computer Science}, 50\penalty0 (1):\penalty0
  1--102, 1987.

\bibitem[Goldstein(1994)]{goldstein-tr94}
S.~C. Goldstein.
\newblock The implementation of a threaded abstract machine.
\newblock Technical Report UCB/CSD-94-818, EECS Department, University of
  California, Berkeley, 1994.
\newblock URL \url{http://www.cs.cmu.edu/~seth/papers/goldstein-tr94.pdf}.

\bibitem[Gonzalez et~al.(2009)Gonzalez, Low, and
  Guestrin]{Gonzalez+al:aistats09paraml}
J.~Gonzalez, Y.~Low, and C.~Guestrin.
\newblock Residual splash for optimally parallelizing belief propagation.
\newblock In \emph{Artificial Intelligence and Statistics (AISTATS)}, 2009.

\bibitem[Gupta et~al.(2001)Gupta, Pontelli, Ali, Carlsson, and
  Hermenegildo]{Gupta:2001:PEP:504083.504085}
G.~Gupta, E.~Pontelli, K.~A.~M. Ali, M.~Carlsson, and M.~V. Hermenegildo.
\newblock Parallel execution of prolog programs: {A} survey.
\newblock \emph{ACM Transactions on Programming Languages and Systems
  (TOPLAS)}, 23\penalty0 (4):\penalty0 472--602, 2001.

\bibitem[Hermenegildo(1986)]{Hermenegildo:1986:AMB:913061}
M.~V. Hermenegildo.
\newblock \emph{An Abstract Machine Based Execution Model for Computer
  Architecture Design and Efficient Implementation of Logic Programs in
  Parallel}.
\newblock PhD thesis, 1986.
\newblock AAI8700203.

\bibitem[Hodas and Miller(1994)]{Hodas94logicprogramming}
J.~S. Hodas and D.~Miller.
\newblock Logic programming in a fragment of intuitionistic linear logic.
\newblock \emph{Information and Computation}, 110:\penalty0 32--42, 1994.

\bibitem[Holzbaur et~al.(2004)Holzbaur, de~la Banda, Stuckey, and
  Duck]{DBLP:journals/corr/cs-PL-0408025}
C.~Holzbaur, M.~J.~G. de~la Banda, P.~J. Stuckey, and G.~J. Duck.
\newblock Optimizing compilation of constraint handling rules in hal.
\newblock \emph{CoRR}, cs.PL/0408025, 2004.

\bibitem[Isard et~al.(2007)Isard, Budiu, Yu, Birrell, and
  Fetterly]{Isard:2007:DDD:1272996.1273005}
M.~Isard, M.~Budiu, Y.~Yu, A.~Birrell, and D.~Fetterly.
\newblock Dryad: distributed data-parallel programs from sequential building
  blocks.
\newblock In \emph{European Conference on Computer Systems (EuroSys)}, pages
  59--72, 2007.

\bibitem[Kacsuk(1990)]{Kacsuk:1990:EMP:533578}
P.~Kacsuk.
\newblock \emph{Execution Models of PROLOG for Parallel Computers}.
\newblock MIT Press, Cambridge, MA, USA, 1990.
\newblock ISBN 0262111497.

\bibitem[Kara et~al.(1997)Kara, Davy, Goodeve, and Nash]{Kara:1997:AMM:265274}
M.~Kara, J.~R. Davy, D.~Goodeve, and J.~Nash, editors.
\newblock \emph{Abstract Machine Models for Parallel and Distributed
  Computing}.
\newblock IOS Press, Amsterdam, The Netherlands, The Netherlands, 1997.
\newblock ISBN 90-5199-267-X.

\bibitem[Lam and Sulzmann(2010)]{DBLP:journals/corr/abs-1006-3039}
E.~S.~L. Lam and M.~Sulzmann.
\newblock Concurrent goal-based execution of constraint handling rules.
\newblock \emph{CoRR}, abs/1006.3039, 2010.

\bibitem[Lin and Kumar(1988)]{Lin:1988:AEL:900478}
Y.~Lin and V.~Kumar.
\newblock And-parallel execution of logic programs on a shared memory
  multiprocessor:a summary of results.
\newblock Technical report, Austin, TX, USA, 1988.

\bibitem[Liu(1998)]{Liu98extendingdatalog}
M.~Liu.
\newblock Extending datalog with declarative updates.
\newblock In \emph{International Conference on Database and Expert System
  Applications (DEXA)}, volume 1873, pages 752--763, 1998.

\bibitem[Loo et~al.(2006)Loo, Condie, Garofalakis, Gay, and
  Hellerstein]{Loo-condie-garofalakis-p2}
B.~T. Loo, T.~Condie, M.~Garofalakis, D.~E. Gay, and J.~M. Hellerstein.
\newblock Declarative networking: Language, execution and optimization.
\newblock In \emph{International Conference on Management of Data (SIGMOD)},
  pages 97--108, 2006.

\bibitem[L\'{o}pez et~al.(2005)L\'{o}pez, Pfenning, Polakow, and
  Watkins]{Lopez:2005:MCL:1069774.1069778}
P.~L\'{o}pez, F.~Pfenning, J.~Polakow, and K.~Watkins.
\newblock Monadic concurrent linear logic programming.
\newblock In \emph{Proceedings of the 7th ACM SIGPLAN International Conference
  on Principles and Practice of Declarative Programming}, PPDP '05, pages
  35--46, New York, NY, USA, 2005.

\bibitem[Low et~al.(2010)Low, Gonzalez, Kyrola, Bickson, Guestrin, and
  Hellerstein]{GraphLab2010}
Y.~Low, J.~Gonzalez, A.~Kyrola, D.~Bickson, C.~Guestrin, and J.~M. Hellerstein.
\newblock Graphlab: {A} new framework for parallel machine learning.
\newblock In \emph{Conference on Uncertainty in Artificial Intelligence (UAI)},
  pages 340--349, 2010.

\bibitem[Malewicz et~al.(2010)Malewicz, Austern, Bik, Dehnert, Horn, Leiser,
  and Czajkowski]{Malewicz:2010:PSL:1807167.1807184}
G.~Malewicz, M.~H. Austern, A.~J.~C. Bik, J.~C. Dehnert, I.~Horn, N.~Leiser,
  and G.~Czajkowski.
\newblock Pregel: a system for large-scale graph processing.
\newblock In \emph{International Conference on Management of Data (SIGMOD)},
  pages 135--146, 2010.

\bibitem[Miller(1985)]{Miller85anoverview}
D.~Miller.
\newblock An overview of linear logic programming.
\newblock In \emph{in Computational Logic}, pages 1--5, 1985.

\bibitem[Ramakrishnan and Ullman(1993)]{Ramakrishnan93asurvey}
R.~Ramakrishnan and J.~D. Ullman.
\newblock A survey of research on deductive database systems.
\newblock \emph{Journal of Logic Programming}, 23:\penalty0 125--149, 1993.

\bibitem[Sunderam(1990)]{Sunderam90pvm:a}
V.~S. Sunderam.
\newblock Pvm: A framework for parallel distributed computing.
\newblock \emph{Concurrency: Practice and Experience}, 2:\penalty0 315--339,
  1990.

\bibitem[{van Rossum}(1995)]{vanRossum:1995:PRM}
G.~{van Rossum}.
\newblock {Python} reference manual.
\newblock Report CS-R9525, Centrum voor Wiskunde en Informatica, Amsterdam, the
  Netherlands, Apr. 1995.
\newblock URL \url{http://www.python.org/doc/ref/ref-1.html}.

\bibitem[Warren(1983)]{AICPub641:1983}
D.~H.~D. Warren.
\newblock An abstract prolog instruction set.
\newblock Technical Report 309, AI Center, SRI International, 333 Ravenswood
  Ave., Menlo Park, CA 94025, Oct 1983.

\bibitem[Warren(1987)]{Warren:1987:OEM:67683.67699}
D.~H.~D. Warren.
\newblock Or-parallel execution models of prolog.
\newblock In \emph{II and Colloquium on Functional and Logic Programming and
  Specifications (CFLP) on TAPSOFT '87: Advanced Seminar on Foundations of
  Innovative Software Development}, pages 243--259, New York, NY, USA, 1987.
  Springer-Verlag New York, Inc.
\newblock ISBN 0-387-17611-X.
\newblock URL \url{http://dl.acm.org/citation.cfm?id=67683.67699}.

\bibitem[Watkins et~al.(2004)Watkins, Cervesato, Pfenning, and
  Walker]{Watkins:2004uq}
K.~Watkins, I.~Cervesato, F.~Pfenning, and D.~Walker.
\newblock A concurrent logical framework: The propositional fragment.
\newblock In \emph{Types for Proofs and Programs}, volume 3085 of \emph{Lecture
  Notes in Computer Science}, pages 355--377. 2004.
\newblock \doi{10.1007/978-3-540-24849-1_23}.

\end{thebibliography}

\end{document}
